### 2011-05-09 bib/prooftransform/redirecting-proofs-2/paper.tex
added draft of tech report explaining one approach to redirect proofs by contradiction

\documentclass[withtimes,a4paper,12pt]{easychair}
\RequirePackage{latexsym}
\RequirePackage{empheq}
\RequirePackage{mathptmx}
\usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{amsthm}
\usepackage[english]{babel}
\usepackage{bussproofs}
\usepackage{cite}
\usepackage{color}
\usepackage{multicol}
\usepackage{subfigure}
%\usepackage{MnSymbol}
%\usepackage[scaled=.86]{helvet}
\usepackage[scaled=.83]{beramono}

\newcommand\blkl{\makebox[1em][l]{\{}}
\newcommand\blkr{\makebox[1em][r]{\}}}

\newdimen\bls
\bls=\baselineskip
\newdimen\forhave
\forhave=.8em %% TYPESETTING

\newcommand\preprf{%
  \vskip\abovedisplayskip
  \noindent
  \begin{tabular}{@{}p{.5\textwidth}@{}p{.5\textwidth}@{}}}
\newcommand\postprf{\end{tabular}\par\vskip\belowdisplayskip}
\def\frm#1\hav#2\end{\hfill \ensuremath{#1\kern\forhave} & \ensuremath{\kern-\forhave{} \have #2} \hfill \\[.5\smallskipamount]}
\def\frx#1\hen#2\end{\hfill \ensuremath{#1\kern\forhave} & \ensuremath{\kern-\forhave{} \hence #2} \hfill \\[.5\smallskipamount]}
\def\cases#1{\\[-.6\bls] %% TYPESETTING, -.6 twice
\multicolumn{2}{c}{$\left[#1\right]$}\\[-.6\bls] \\[1.5\smallskipamount]}

% for debugging
\newcommand\HR{\rule{1em}{0.4pt}}
\newcommand\bluefbox[1]{\textcolor{blue}{%
  \setlength\fboxsep{0pt}\fbox{\textcolor{black}{#1}}}}

\newcommand\keyw[1]{\textsf{\textbf{#1}}}
\newcommand{\easychair}{\sf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}

\newcommand\vthinspace{\kern+0.083333em}
\newcommand\negvthinspace{\kern-0.083333em}

\newcommand\smallbot{\kern-.125em\bot\kern-.125em}
\let\B=\overline
%\newcommand\have{\mathrel{\vdash}}
%\newcommand\have{\mathrel{\;\hookrightarrow\;}}
\newcommand\have{\mathrel{\;\vartriangleright\;}}
\newcommand\hence{\mathrel{\;\blacktriangleright\;}}
%\newcommand\have{\mathrel{\,\rightharpoonup\,}}
\newcommand\havebot{\have\smallbot} %% TYPESETTING
\newcommand\hencebot{\hence\smallbot} %% TYPESETTING

\EnableBpAbbreviations
\def\ScoreOverhang{2pt}
\def\proofSkipAmount{\vskip 0pt}
\def\defaultHypSeparation{\hskip0.75em}

\begin{document}

%\title{Redirecting Proofs by Contradiction}
\title{Redirecting Resolution Proofs}
\titlerunning{Redirecting Resolution Proofs}

%\volumeinfo
%	{G. Sutcliffe, A. Voronkov}         % editors
%	{2}                                 % number of editors
%	{{\easychair} 2.0 Beta 4, October 2010}      % event
%	{2}                                 % volume
%	{1}                                 % issue
%	{1}                                 % starting page number

%\headfootstyle
%	{}
%	{\sf}
%	{\footnotesize\sf}
%	{\small}

%\volumeinfoECPS
%	{EasyChair Workshop Proceedings}
%	{ECPS vol. 7999}

\author{
    Jasmin Christian Blanchette\\
    \affiliation{Technische Universit\"at M\"unchen, Germany}\\
    \affiliation{\url{blanchette@in.tum.de}}\\
}

\authorrunning{J. C. Blanchette}

\maketitle

%\begin{abstract}
%\noindent
%  * ATP proofs are hard to read
%  * some big frameworks
%  * here we present a simple algo to redirect proofs by contradiction
%\end{abstract}

\section{Introduction}
\label{sec:introduction}

The proofs returned by automatic theorem provers (ATPs) are notoriously
difficult to read. This is an issue if an ATP solves an open mathematical
problem (such as the Robbins conjecture \cite{mccune-1997}), because users then
certainly want to study the proof closely. But even in the context of program
verification, where users are normally satisfied with a ``proved'' or
``disproved,'' they might still want to look at the proofs---for example, if
they suspect that their axioms are inconsistent.

Our interest in readable ATP proofs has a different origin. The tool
Sledgehammer \cite{meng-paulson-2008,paulson-susanto-2007} integrates
ATPs with the interactive theorem prover Isabelle/HOL
\cite{nipkow-et-al-2002}. Given an Isabelle/HOL conjecture,
Sledgehammer heuristically selects a few hundred relevant lemmas from
Isabelle's libraries, translates them along with the conjecture, and
sends the resulting problem to E~\cite{schulz-2004},
SPASS~\cite{weidenbach-2001}, and
Vampire~\cite{riazanov-voronkov-2002}.
%
Proofs are reconstructed in Isabelle either using a single invocation
of the built-in resolution prover Metis \cite{hurd-2003} or as
structured Isar \cite{wenzel-2007} proofs following the ATP proofs
\cite{paulson-susanto-2007}. This latter option is useful for larger
proofs, which Metis fails to re-find within a reasonable time. But
most users find the proofs hideous and are little enclined to
insert them in their formalizations.

As illustration, consider the conjecture ``$\textit{length}~(\textit{tl}~\textit{xs}) \le
\textit{length~xs}$'', which states that the tail of a list (the list from which
we remove its first element, or the empty list if the list is empty) is shorter
than or of equal length as the original list. The proof produced by Vampire,
translated to Isabelle's structured Isar format, looks as follows:
%
\begin{quote}
\keyw{proof}\,~\textit{neg\_clausify} \\
\hbox{}\quad \keyw{assume}\, ``$\lnot\;\, \textit{length}~(\textit{tl}~\textit{xs}) \le \textit{length}~\textit{xs}$''\\
\hbox{}\quad \keyw{hence}\, ``$\textit{drop}~(\textit{length}~\textit{xs})~(\textit{tl}~\textit{xs}) \not= \text{[]}$'' \keyw{\,by\,} (\textit{metis\, drop\_eq\_Nil\/})\\
\hbox{}\quad \keyw{hence}\, ``$\textit{tl}~(\textit{drop}~(\textit{length}~\textit{xs})~\textit{xs}) \not= \text{[]}$'' \keyw{\,by\,} (\textit{metis\, drop\_tl\/})\\
\hbox{}\quad \keyw{hence}\, ``$\forall\textit{u}.\;\, \textit{xs} \mathbin{@} \textit{u} \not= \textit{xs} \mathrel{\lor} \textit{tl}~\textit{u} \not= \text{[]}$'' \keyw{\,by\,} (\textit{metis\, append\_eq\_conv\_conj\/})\\
\hbox{}\quad \keyw{hence}\, ``$\textit{tl}~\text{[]} \not= \text{[]}$'' \keyw{\,by\,} (\textit{metis\, append\_Nil2\/})\\
\hbox{}\quad \keyw{thus}\, ``$\textit{False}$'' \keyw{\,by\,} (\textit{metis}\,~\textit{tl.simps\/}(1))\\
\keyw{qed}
\end{quote}
%
\noindent
The \textit{neg\_clausify} method transforms the Isabelle conjecture into
negated clause form, ensuring that it has the same shape as the corresponding ATP
conjecture. The negation of the clause is introduced by the
\keyw{assume} keyword, and a series of intermediate facts introduced by
\keyw{hence} lead to a contradiction.

The first obstacle for readability is that the Isar proof, like the
underlying ATP proof, is by contradiction. Such proofs can be turned
around by applying contraposition repeatedly. For example, the above
proof can be transformed into
%
\begin{quote}
\keyw{proof} -- \\
\hbox{}\quad \keyw{have}\, ``$\textit{tl}~\text{[]} = \text{[]}$'' \keyw{\,by\,} (\textit{metis}\,~\textit{tl.simps\/}(1)) \\
\hbox{}\quad \keyw{hence}\, ``$\exists\textit{u}.\;\, \textit{xs} \mathbin{@} \textit{u} = \textit{xs} \mathrel{\land} \textit{tl}~\textit{u} = \text{[]}$'' \keyw{\,by\,} (\textit{metis\, append\_Nil2}) \\
\hbox{}\quad \keyw{hence}\, ``$\textit{tl}~(\textit{drop}~(\textit{length}~\textit{xs})~\textit{xs}) = \text{[]}$'' \keyw{\,by\,} (\textit{metis\, append\_eq\_conv\_conj\/}) \\
\hbox{}\quad \keyw{hence}\, ``$\textit{drop}~(\textit{length}~\textit{xs})~(\textit{tl}~\textit{xs}) = \text{[]}$'' \keyw{\,by\,} (\textit{metis\, drop\_tl\/}) \\
\hbox{}\quad \keyw{thus}\, ``$\textit{length}~(\textit{tl}~\textit{xs}) \le \textit{length}~\textit{xs}$'' \keyw{\,by\,} (\textit{metis\, drop\_eq\_Nil\/}) \\
\keyw{qed}
\end{quote}
%
Most Isabelle users find the direct proof much easier to understand
and maintain. The proof is still somewhat difficult to read because
lemmas are referred to by name, but it becomes clear once we reveal
them:
\[\hbox{\begin{tabular}{@{}r@{\kern.75em}l@{}}%
\textit{tl.simps\/}(1): & $\textit{tl}~\text{[]} = \text{[]}$ \\
\textit{append\_Nil2\/}: & $\textit{xs} \mathbin{@} \text{[]} = \textit{xs}$ \\
\textit{append\_eq\_conv\_conj\/}: & $\textit{xs} \mathbin{@} \textit{ys} = \textit{zs} \,\longleftrightarrow\,
\textit{xs} = \textit{take}~(\textit{length}~\textit{xs})~\textit{zs} \mathrel{\land}
\textit{ys} = \textit{drop}~(\textit{length}~\textit{xs})~\textit{zs}$ \\
\textit{drop\_tl\/}: & $\textit{drop}~n~(\textit{tl}~\textit{xs}) = \textit{tl}~(\textit{drop}~n~\textit{xs})$ \\
\textit{drop\_eq\_Nil\/}: & $\textit{drop}~n~\textit{xs} = \text{[]} \,\longleftrightarrow\, \textit{length}~\textit{xs} \leq n$
\end{tabular}}\]
The direct proof also forms a good basis for further development: The
user can clean it up further to increase readability and
maintainability.

There is a large body of research about making resolution proofs
readable. Earlier work focused on translating detailed resolution
proofs into natural deduction or sequent calculi \cite{xxx, etc.}.
Although they are arguably more readable, these calculi still operate
at the logical level, whereas humans reason mostly at the assertion
level, invoking definitions and lemmas without providing the full
logical details \cite{huang-xxx}. A line of research focused on
transforming natural deduction proofs into assertion-level proofs,
culminating with the systems \textsc{Tramp} \cite{xxx} and Otterfier
\cite{zimmer-et-al-2004}.

We would have liked to try out \textsc{Tramp} and Otterfier, but these
are large pieces of unmaintained software that are hardly installable
on modern machines and that only support older ATPs. Regardless, if we
look at Sledgehammer proofs, the problem looks somewhat different.
Modern ATPs produce proofs with fairly large steps. Because
Sledgehammer supplies the ATPs with hundreds of lemmas, they tend to
find short proofs, typically involving a handful of lemmas. Moreover,
Sledgehammer merges consecutive logical steps into larger steps and
can be further instructed to keep only each $n$th larger step, if
short proofs are desired. Replaying is also not such an important
issue for us, since we can rely on the fairly powerful Metis prover.

As a first step toward more intellible proofs, I looked for a method
to turn contradiction proofs around. It seemed obvious the
method should not be tied to any logic as long as it is classical, or
to any calculus. In particular, it should work on the Isar proofs
generated by Sledgehammer, or directly on first-order TSTP proofs.
Finally, the direct proof should be expressible comfortably in a
block-structured Isar-like language, with case splits and nested
subproofs.

In my PAAR 2010 and IWIL 2010 talks \cite{xxx,xxx}, I sketched a
method for transforming proofs by contradiction into direct proofs.
The method was very simple and simply exploited the contrapositive,
but in the worst case the resulting proof could explode exponentially
in size. In this paper, I propose a new method transform resolution
proofs into direct proofs expressed in a simple Isar-like syntax
(Section~\ref{sec:proof-notations}), based on a simple set of rules
(Sections \ref{sec:introductory-examples} and
\ref{sec:the-algorithm}). The new approach is sound and complete and
has the advantage that each deduction in the ATP proof
gives rise to exactly one deduction in the direct proof. A linear
number of additional steps are introduced in the direct proof, but
these are simple logical rules, such as modus ponens, and do not
require the full power of Metis.
%The approach is illustrated on
%actual ATP proofs found by Sledgehammer (Section~\ref{sec:gallery}).

\section{Proof Notations}
\label{sec:proof-notations}

%\def\epiind{\noindent\hbox{}\kern.4\textwidth}
%\epiind ``A proof is a proof. What kind of a proof? It's a proof. \\
%\epiind A~proof is a proof, and when you have a good proof, \\
%\epiind it's because it's proven.'' \\
%\epiind \qquad --- Jean Chr\'etien (2002)
%\medskip

First, we need to define the format we want to use for proofs. We need
several formats.

\paragraph{Proof graph.} A proof graph is a directed acyclic graph where an
arrow $a \rightarrow b$ indicates that $a$ is used to derive $b$. By ensuring
that derived nodes appear lower than their parent nodes in the graph, we can
omit the arrowheads:
%
\[\hbox{\includegraphics[scale=0.45]{diagrams/paar-ref-graph.pdf}}\]
%
ATP proofs usually identify formulas (typically clauses) by numbers: for
example, the conjecture might be called 0, the axioms might be numbered 1, 2,
\ldots, 250, and new derivations might be called 251 and above. In this paper,
we abstract the ATP proofs by ignoring the actual formulas and just keeping the
numbers. Also, for our own convenience, we put a bar on top of the negated
conjecture and all the formulas derived directly or indirectly from it, denoting
negation. To unnegate the conjecture, or negate any of the steps that are
\emph{tainted} by the negated conjecture, we simply remove the bar. For the last
step, we write $\bot$ rather than $\B{\top}$.

\paragraph{Proof trees.} Proof trees are the standard notation for natural
deduction proofs. For example, the proof graph above is captured by the
following proof tree:
%
\begin{prooftree}
\AXC{$[\B{0}]$}
\UIC{$\B{0}$}
    \AXC{$1$}
        \AXC{$2$}
    \BIC{$1 \mathrel{\land} 2$}
        \AXC{$1 \mathrel{\land} 2 \longrightarrow 4$}
    \BIC{$4$}
\BIC{$\B{0} \mathrel{\land} 4$}
    \AXC{$\kern-4em\B{0} \mathrel{\land} 4 \longrightarrow \B{6}$}
\BIC{$\B{6}$}
    \AXC{$[\B{0}]$}
    \UIC{$\B{0}$}
        \AXC{$3$}
            \AXC{$1$}
                \AXC{$2$}
            \BIC{$1 \mathrel{\land} 2$}
                \AXC{$1 \mathrel{\land} 2 \longrightarrow 4$}
            \BIC{$4$}
        \BIC{$3 \mathrel{\land} 4$}
            \AXC{$\kern-3em 3 \mathrel{\land} 4 \longrightarrow 5$}
        \BIC{$5$}
    \BIC{$\B{0} \mathrel{\land} 5$}
        \AXC{$\kern-5em\B{0} \mathrel{\land} 5 \longrightarrow \B{7}$}
    \BIC{$\B{7}$}
\BIC{$\B{6} \mathrel{\land} \B{7}$}
    \AXC{$\kern-5em\B{6} \mathrel{\land} \B{7} \longrightarrow \bot$}
\BIC{$\bot$}
\end{prooftree}
The proof graph notation is more compact, not least because it enables
sharing of subproofs. For that reason, we will prefer them to proof trees.

\paragraph{Isar proofs.} Isar proofs are a linearization of natural deduction
proofs, but unlike proof trees they enable sharing:
%
\begin{quote}
\keyw{proof}\,~\textit{neg\_clausify} \\
\hbox{}\quad \keyw{assume}\, $\B{0}$ \\
\hbox{}\quad \keyw{have}\, $4$ \keyw{\,by\,} (\textit{metis\, }$1$ $2$) \\
\hbox{}\quad \keyw{have}\, $5$ \keyw{\,by\,} (\textit{metis\, }$3$ $4$) \\
\hbox{}\quad \keyw{have}\, $\B{6}$ \keyw{\,by\,} (\textit{metis\, }$\B{0}$ $4$) \\
\hbox{}\quad \keyw{have}\, $\B{7}$ \keyw{\,by\,} (\textit{metis\, }$\B{0}$ $5$) \\
\hbox{}\quad \keyw{show} \kern.03em$\bot$ \keyw{\kern.05em by\,} (\textit{metis\, }$\B{6}$ $\B{7}$) \\ %% TYPESETTING
\keyw{qed}
\end{quote}
%
The above proof is by contradiction. A direct proof of the above in Isar would
be
%
\begin{quote}
\keyw{proof} -- \\
\hbox{}\quad \keyw{have}\, $4$ \keyw{\,by\,} (\textit{metis\, }$1$ $2$) \\
\hbox{}\quad \keyw{have}\, $5$ \keyw{\,by\,} (\textit{metis\, }$3$ $4$) \\
\hbox{}\quad \keyw{have}\, $6 \mathrel{\lor} 7$ \keyw{\,by\,} \textit{metis} \\
\hbox{}\quad $\{$\quad \keyw{assume}\, $6$ \\
\hbox{}\qquad\phantom{$\{$}\keyw{have}\, $0$ \keyw{\,by\,} (\textit{metis\, }$4$ $6$)\quad$\}$ \\
\hbox{}\quad \keyw{moreover} \\
\hbox{}\quad $\{$\quad \keyw{assume}\, $7$ \\
\hbox{}\qquad\phantom{$\{$}\keyw{have}\, $0$ \keyw{\,by\,} (\textit{metis\, }$5$ $7$)\quad$\}$ \\
\hbox{}\quad \keyw{ultimately show}\, $0$ \keyw{\,by\,} (\textit{metis\, }`$6 \mathrel{\lor} 7$') \\
\keyw{qed}
\end{quote}
We refer to the Isar tutorial \cite{nipkow-2011-isar} for more information on
the Isar syntax.

\paragraph{Shorthand proofs.}
The last proof format we need to review is a shorthand notation for a subset of
Isar proofs. In their simplest form, these shorthand proofs are simply a list of
derivations, where `$a_1, \ldots, a_n \have c$' means ``from $a_1$ and $\ldots$
and $a_n$, we conclude $c$.'' If among the derivation's assumptions $a_j$
we have the previous derivation's conclusion, we can omit the assumption
and write `$a_1, \ldots, a_{j-1}, a_{j+1}, \ldots, a_n \hence c$' (corresponding
to Isar's \keyw{hence} and \keyw{thus} keywords). Depending on whether we use
the abbreviated format, our running example becomes
%
\begin{align*}
1, 2 & \have 4 &
    1, 2 & \have 4 \\
3, 4 & \have 5 &
    3 & \hence 5 \\
\B{0}, 4 & \have \B{6} &
    \B{0}, 4 & \have \B{6} \\
\B{0}, 5 & \have \B{7} &
    \B{0}, 5 & \have \B{7} \\
\B{6}, \B{7} & \have \bot &
    \B{6} & \hence \bot
\end{align*}
%
Each step is essentially a sequent $\Gamma \vdash \varphi$. The assumptions are
either the negated conjecture ($\B{0}$), facts that were proved elsewhere ($1$,
$2$, and $3$), or formulas that were proved in preceding sequents ($4$, $5$,
$\B{6}$, and $\B{7}$). The succedent of the last sequent is empty ($\bot$).

Direct proofs can be done the same way, but they have the constraints that we
may not assume the negated conjecture $\B{0}$ in any of the sequent and that the
last sequent has the conjecture $0$ as succedent. However, in some of the direct
proofs, we will find it useful to introduce case splits, as we had in the Isar
proof above. For example:
%
\[\hbox{\begin{tabular}{@{}p{.5\textwidth}@{}p{.5\textwidth}@{}}
\frm 1, 2 \hav 4 \end
\frx 3 \hen 5 \end
\frm \hav 6, 7 \end
\cases{\begin{tabular}{l|l}
    $\begin{aligned}[t]
      [&6] \\
      4 \hence {} & 0
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t] 
      [&7] \\
      5 \hence {} & 0
    \end{aligned}$
  \end{tabular}}
\end{tabular}}\]
The notation in brace is a case split. In general, a case split has the form
\[\left[\begin{tabular}{l|c|l}
  $\begin{aligned}[t]
    [&\Gamma_1] \\
     \Delta_{11} \have {} & \mathrm{E}_{11} \\
     & ~\vdots \\
     \Delta_{1k_1} \have {} & \mathrm{E}_{1k_1}
    \end{aligned}$\enskip\hbox{}&\enskip
  $\begin{aligned}[t]
    ~\ldots~ \\
    ~\ldots~ \\
	\\
    ~\ldots~
    \end{aligned}$\enskip\hbox{}&\enskip
  $\begin{aligned}[t]
    [&\Gamma_n] \\
     \Delta_{n1} \have {} & \mathrm{E}_{n1} \\
     & ~\vdots \\
     \Delta_{nk_n} \have {} & \mathrm{E}_{nk_n}
    \end{aligned}$
\end{tabular}
\right]\]
with the requirement that a sequent with the succedent $\Gamma_1,
\ldots, \Gamma_n$ has been proved already. Also, each of the branches
must be a valid proof. The assumptions $[\Gamma_{\!j}]$ may be used to
discharge assumptions in the same branch, just as if they had been
sequents ${} \have \Gamma_{\!j}$. Seen from outside, the case split
expression stands for a sequent with the succedent
$\mathrm{E}_{1k_1}, \ldots, \mathrm{E}_{nk_n}$.

\section{Introductory Examples}
\label{sec:introductory-examples}

\subsection{A Linear Proof}
\label{sec:a-linear-proof}

We start with a simple proof by contradiction expressed as a proof tree and
using our shorthand notation:
%
\[\hbox{\begin{tabular}{l@{\qquad}l}
\raisebox{-\height}{\includegraphics[scale=0.45]{diagrams/linear-ref-graph.pdf}}
&
\qquad
\raisebox{-\height}{$\begin{aligned}[t]
\\
\B 0, 1 & \have \B 3 \\
2, \B 3 & \have \B 4 \\
1, \B 4 & \havebot
\end{aligned}$}
\end{tabular}}\]
Recall from Sect.\ \ref{sec:proof-notations} that the negated conjecture all all
the proof steps that are tainted by it are shown as negated. Next, we turn the
sequents around using contraposition at the sequent level to avoid all
negations. This gives
\begin{align*}
1, 3 & \have 0 \\
2, 4 & \have 3 \\
1 & \have 4
\end{align*}
Finally, if we reorder the sequents and introduce $\hence$ wherever possible, we
obtain a direct proof:

\begin{center}
\begin{tabular}{@{}p{.3\textwidth}@{}p{.3\textwidth}@{}}
\begin{tabular}{@{}l@{}}
\keyw{proof} -- \\
\hbox{}\quad \keyw{have}\, $4$ \keyw{\,by\,} \textit{metis} \\
\hbox{}\quad \keyw{hence}\, $3$ \keyw{\,by\,} (\textit{metis\, }$2$) \\
\hbox{}\quad \keyw{thus}\, $0$ \keyw{\,by\,} (\textit{metis\, }$1$) \\
\keyw{qed}
\end{tabular}
&
\begin{tabular}{@{}p{.25\textwidth}@{}p{.25\textwidth}@{}}
\frm \hav 4 \end
\frx 2 \hen 3 \end
\frx 1 \hen 0 \end
\end{tabular}
\end{tabular}
\end{center}

\subsection{Two Lasso-Shaped Proofs}
\label{sec:two-lasso-shaped-proofs}

The next two examples are lasso-shaped proofs:
%
\[\hbox{\begin{tabular}{llll}
\raisebox{-\height}{\includegraphics[scale=0.45]{diagrams/lasso-1-ref-graph.pdf}}
&
\qquad
\raisebox{-\height}{$\begin{aligned}[t]
\\
\B 0 & \have \B 1 \\
\B 0 & \have \B 2 \\
\B 1 & \have \B 3 \\
\B 2 & \have \B 4 \\
\B 3, \B 4 & \have \B 5 \\
\B 5 & \have \B 6 \\
\B 6 & \havebot
\end{aligned}$}
&
\qquad\qquad
\raisebox{-\height}{\includegraphics[scale=0.45]{diagrams/lasso-2-ref-graph.pdf}}
&
\qquad
\raisebox{-\height}{$\begin{aligned}[t]
\\
\B 0 & \have \B 1 \\
\B 1 & \have \B 2 \\
\B 2 & \have \B 3 \\
\B 2 & \have \B 4 \\
\B 3 & \have \B 5 \\
\B 4 & \have \B 6 \\
\B 5, \B 6 & \havebot
\end{aligned}$}
\end{tabular}}\]

We start with the contradiction proof on the left-hand side. Starting from
$\bot$, it is easy to turn the proof around up to the lasso cycle:
%
\begin{align*}
& \have 6 \\
& \hence 5 \\
& \hence 3 \lor 4
\end{align*}
%
When applying the contrapositive to eliminate the negations in $\B 3, \B 4 \have
\B 5$, we obtain a disjunction on the right-hand side of the sequent: $5 \have 3
\lor 4$. To continue from there, we need a case split. Then we can finish then
branches:
%
\[\left[\begin{tabular}{l|l} %% TYPESETTING
    $\begin{aligned}[t]
      [&3] \\
      \hence {} & 1 \\
      \hence {} & 0
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t] 
      [&4] \\
      \hence {} & 2 \\
      \hence {} & 0
    \end{aligned}$
  \end{tabular}\right]\]

The second lasso example is more tricky, because the cycle occurs near the end
of the contradiction proof. Already when redirecting the last inference, we get
a disjunction:
%
\begin{align*}
& \have 5 \lor 6
\end{align*}
%
And if we split naively and finish each branch independently of each other,
we end up with a fair share of duplication:
\pagebreak[3]
\[\left[\begin{tabular}{l|l} %% TYPESETTING
    $\begin{aligned}[t]
      [&5] \\
      5 \have {} & 3 \\
      3 \have {} & 2 \\
      2 \have {} & 1 \\
      1 \have {} & 0 \\
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t] 
      [&6] \\
      6 \have {} & 4 \\
      4 \have {} & 2 \\
      2 \have {} & 1 \\
      1 \have {} & 0 \\
    \end{aligned}$
  \end{tabular}\right]\]

In this case, the key is to observe that it is enough if both branches prove
$2$, and from there we prove the rest. The complete proof is (without and
with $\hence$):
\[\hbox{\begin{tabular}{@{}p{.25\textwidth}@{}p{.25\textwidth}@{}}
\frm \hav 5 \lor 6 \end
\cases{\begin{tabular}{l|l}
    $\begin{aligned}[t]
      [&5] \\
      5 \have {} & 3 \\
      3 \have {} & 2
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t] 
      [&6] \\
      6 \have {} & 4 \\
      4 \have {} & 2
    \end{aligned}$
  \end{tabular}}
\frm 2 \hav 1 \end
\frm 1 \hav 0 \end
\end{tabular}\begin{tabular}{@{}p{.25\textwidth}@{}p{.25\textwidth}@{}}
\frm \hav 5 \lor 6 \end
\cases{\begin{tabular}{l|l}
    $\begin{aligned}[t]
      [&5] \\
      \hence {} & 3 \\
      \hence {} & 2
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t] 
      [&6] \\
      \hence {} & 4 \\
      \hence {} & 2
    \end{aligned}$
  \end{tabular}}
\frx 2 \hen 1 \end
\frx \hen 0 \end
\end{tabular}}\]

%Refer to $\eta$ rule or whatever.

Here we were lucky because we could join both branches the node 2 to avoid all
duplication. If we want to stick to a strick no-duplication policy, in general
we sometimes need to join on a disjunction $\varphi_1 \lor \cdots \lor
\varphi_n$, as the next example will illustrate.

\subsection{A Diabolical Proof}
\label{sec:a-diabolical-proof}

Our final example is truly diabolical (and sightly unrealistic):
%
\[\hbox{\begin{tabular}{llll}
\raisebox{-\height}{\includegraphics[scale=0.45]{diagrams/diabolical-ref-graph.pdf}}
&
\qquad
\raisebox{-\height}{$\begin{aligned}[t]
\\
\B 0 & \have \B 1 \\
\B 1 & \have \B 2 \\
\B 1 & \have \B 3 \\
\B 2, \B 3 & \have \B 4 \\
\B 2, \B 4 & \have \B 5 \\
\B 3, \B 4 & \have \B 6 \\
\B 2, \B 5, \B 6 & \have \B 7 \\
\B 3, \B 6 & \have \B 8 \\
\B 7, \B 8 & \havebot
\end{aligned}$}
\end{tabular}}\]

We start with

$      |- 7 \lor 8$

Then we identify the nodes that dominate only one branch in the refutation
graph---i.e., that are reachable by navigating backward (upward). Looking at the
graph, we see that $\B 5$ dominates $\B 7$ but not $\bot 8$, and is the only
such node. So we do a case split, and allow ourselves to perform inferences
requiring 5 and 7 in the left branch and 8 in the right branch. First the
right branch:

    [8]
    8 |- 3, 6

Nothing more to do. Any further inferences we would do on it
would need to be repeated in the left branch, which we want to avoid.
Left branch:

         [7]
$    7 |- 2 \lor 5 \lor 6$

    { [2]
    | [5]
$      5 |> 2 \lor 4$
    | [6]}

The 2 and 6 are left alone. We can abbreviate the subcase split to

  $2 \lor 5 \lor 6 |- 2 \lor 4 \lor 6$

In the end, the left branch proves $2 \lor 4 \lor 6$, and
both branches together prove $2 \lor 3 \lor 4 \lor 6$.

This gives us a new case split, but notice that 6 is dominated by every other.
We start with it:

  {[2]
  |[3]
  |[4]
  |[6]
$   6 |- 3 \lor 4$}

i.e. $2 \lor 3 \lor 4$. Since all but one branches are trivial, we abbreviate it
as

$  2 \lor 3 \lor 4 \lor 6 |- 2 \lor 3 \lor 4$

One way to see this is as ``rewriting.'' Similarly with 4:

  {[2]
  |[3]
  |[4]
$   4 |- 2 \lor 3$}

abbreviated as

$  2 \lor 3 \lor 4 |- 2 \lor 3$

i.e. $2 \lor 3$. The rest is reminiscent of our second lasso:

  { [2]
    2 |- 1
  | [3]
    3 |- 1 }
  1 |- 0

Gluing all of this together, we get

\begin{quote}
\begin{tabular}{@{}p{.45\textwidth}@{}p{.5\textwidth}@{}}
\begin{tabular}{@{}l@{}}
\keyw{proof} -- \\
\hbox{}\quad \keyw{have}\, $7 \lor 8$ \keyw{\,by\,} \textit{metis} \\
\hbox{}\quad \keyw{moreover} \\
\hbox{}\quad \blkl \keyw{assume}\, $7$ \\
\hbox{}\qquad \keyw{hence}\, $2 \lor 5 \lor 6$ \keyw{\,by\,} \textit{metis} \\
\hbox{}\qquad \keyw{hence}\, $2 \lor 4 \lor 6$ \keyw{\,by\,} \textit{metis}\blkr \\
\hbox{}\quad \keyw{moreover} \\
\hbox{}\quad \blkl \keyw{assume}\, $8$ \\
\hbox{}\qquad \keyw{hence}\, $3 \lor 6$ \keyw{\,by\,} \textit{metis}\blkr \\
\hbox{}\quad \keyw{ultimately have}\, $2 \lor 3 \lor 4 \lor 6$ \keyw{\,by\,} \textit{metis} \\
\hbox{}\quad \keyw{hence}\, $2 \lor 3 \lor 4$ \keyw{\,by\,} \textit{metis} \\
\hbox{}\quad \keyw{hence}\, $2 \lor 3$ \keyw{\,by\,} \textit{metis} \\
\hbox{}\quad \keyw{moreover} \\
\hbox{}\quad \blkl \keyw{assume}\, $2$ \\
\hbox{}\qquad \keyw{hence}\, $1$ \keyw{\,by\,} \textit{metis}\blkr \\
\hbox{}\quad \keyw{moreover} \\
\hbox{}\quad \blkl \keyw{assume}\, $3$ \\
\hbox{}\qquad \keyw{hence}\, $1$ \keyw{\,by\,} \textit{metis}\blkr \\
\hbox{}\quad \keyw{ultimately have}\, $1$ \keyw{\,by\,} \textit{metis} \\
\hbox{}\quad \keyw{thus}\, $0$ \keyw{\,by\,} \textit{metis} \\
\keyw{qed}
\end{tabular}
&
\begin{tabular}{@{}p{.25\textwidth}@{}p{.25\textwidth}@{}}
\frm \hav 7 \lor 8 \end
\cases{\begin{tabular}{l|l}
    $\begin{aligned}[t]
      [&7] \\
      \hence {} & 2 \lor 5 \lor 6 \\
      \hence {} & 2 \lor 4 \lor 6
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t]
      & \\
      [&8] \\
      \hence {} & 3 \lor 6
    \end{aligned}$
  \end{tabular}}
\frx \hen 2 \lor 3 \lor 4 \end
\frx \hen 2 \lor 3 \end
\cases{\begin{tabular}{l|l}
    $\begin{aligned}[t]
      [&2] \\
      \hence {} & 1
    \end{aligned}$\enskip\hbox{}&\enskip
    $\begin{aligned}[t]
      [&3] \\
      \hence {} & 1
    \end{aligned}$\end{tabular}}
\frx \hen 0 \end
\end{tabular}
\end{tabular}
\end{quote}

Which isn't too bad, considering the spaghetti we started with.

\section{The Algorithm}
\label{sec:the-algorithm}

There is one more concept that needs to be explained: that of a
``target''. The target set is a set of formulas whose disjunction is
what we want to prove. For the top-level proof, the target is the
singleton consisting only of the conjecture. For subproofs, the target

Our approach is based on four basic tools: contraposition (on
sequents), cut, case split, and ``rewriting.'' Before we look at the
actual algorithm, it will help to see these tools in action.

Cut (our second tool) is implicit in this notation. The proof
tree makes it explicit:
%
\begin{prooftree}
\AXC{\phantom{.}}
\UIC{${} \have 2$\strut}
  \AXC{\phantom{.}}
  \UIC{$2 \have 1$\strut}
\RL{\textsc{Cut}}
\BIC{${} \have 1$\strut}
  \AXC{\phantom{.}}
  \UIC{$1 \have 0$\strut}
\RL{\textsc{Cut}}
\BIC{${} \have 0$\strut}
\end{prooftree}

In general, reordering might not suffice. For example, for the proof
by contradiction used as the running example in
Section~\ref{sec:proof-notations}, the contrapositive sequents are
%
\begin{align*}
1, 2 & \have 4 \\
3, 4 & \have 5 \\
4, 6 & \have 0 \\
5, 7 & \have 0 \\
{} & \have 6, 7
\end{align*}

We can start outputing the proof
%
\begin{align*}
1, 2 & \have 4 \\
3, 4 & \have 5 \\
{} & \have 6, 7
\end{align*}
%
but then we get stuck: There is no obvious way to use the remaining
two sequents, $4, 6 \have 0$ and $5, 7 \have 0$, because we only know
``$6$ or $7$.'' With a case split (our third tool), we can finish the
proof. In the left branch, we assume $6$, which allows us to use the
sequent $4, 6 \have 0$ to deduce $0$, and similarly in the right
branch. We saw the full proof in Section~\ref{sec:proof-notations}.
Case split can be seen as a special case of a multi-cut rule.

Alternative: Using ``rewriting'' (our fourth tool), the sequent ${}
\have 6, 7$ can be transformed until we reach the desired conclusion.
E.g.
%
\begin{align*}
4 & \have 0, 7 \\
4, 5 & \have 0
\end{align*}

This is justified by the following proof tree:
%
\begin{prooftree}
\AXC{${} \have 6, 7$\strut}
  \AXC{$4, 6 \have 0$\strut}
\RL{\textsc{Cut}}
\BIC{$4 \have 0, 7$\strut}
  \AXC{$5, 7 \have 0$\strut}
\RL{\textsc{Cut}}
\BIC{$4, 5 \have 0$\strut}
\end{prooftree}
%
However, in this case, we prefer the case split, because it more
clearly reflects the structure of the underlying proof. As the proof
tree indicates, rewriting is again an instance of cut.

%\section{Theoretical Properties}
%\label{sec:theoretical-properties}
%
%\subsection{Soundness}
%
%  contrapositive
%  MCut
%
%\subsection{Termination}
%
%\subsection{Linear Complexity}
%
%\section{Gallery}
%\label{sec:gallery}
%
%* soundness and completeness are nice, but what matters is that the
%  results are somehow aesthetically pleasing
%* saw few examples of certain shapes
%* actual proof from \textit{Metis\_Examples} (or elsewhere?)
%
%\subsection{Big O Notation}
%
%
%\subsection{Typed Set Theory}
%
%
%\subsection{Tarski}
%

\section{Conclusion}
\label{sec:conclusion}

  * method is not terribly difficult
  * and if would not surprise us if it was part of folklore or
    some special case of some existing work

In a second step, we are interested in transformations that increase
proof readability, such as those available for Mizar proofs
\cite{pak-2010}.

\paragraph{Acknowledgment.}
Tobias Nipkow.
Geoff Sutcliffe and Joseph Urban motivated me to look into this.
Andrei Popescu, with his deep theoretical insights, helped simplify my
previous framework.

\bibliography{bib}{}
\bibliographystyle{abbrv}

\end{document}